% Abstract and keywords for the IEEE TC Graph-Centric Computing special issue.

\begin{abstract}
Low Earth orbit constellations induce \emph{dynamic graphs}: ground stations and satellites
form a time-varying topology whose edges appear, disappear, and change weight as links
hand over and as paths reoptimize.
Distributed routing protocols and centralized software-defined controllers therefore
face the same graph-centric question: how does a system update routes when the underlying
graph changes, and at what cost to data-plane reachability as graph size grows?

We study this question in StarryNet, a container-based satellite network emulator that
runs full routing stacks under reproducible, geometry-driven topology dynamics.
We compare OSPF (distributed link-state routing) with an SDN controller that performs
proactive, incremental, make-before-break route installation before ground-satellite
link teardown.
Both modes are evaluated with identical workloads, identical probe pairs, and a
continuous data-plane outage metric that measures sustained loss bursts aligned to
control events, rather than tick-scheduled pings that can hide multi-second black holes.

We scale constellations from $5\times 5$ to $10\times 10$ (27 to 102 emulated nodes),
with ten independent repetitions per grid size and routing mode (120 runs total).
At the first handover event in each run, SDN proactive handover yields
0\,ms mean data-plane outage across all six grid sizes
(0\,ms at 102 nodes over ten repetitions).
OSPF exhibits multi-second topology-change outages at the same events
(e.g., 12.9\,s mean at 51 nodes and 5.7\,s at 38 nodes on the first handover),
while SDN maintains 0\% packet loss at the handover and post-handover ping phases
for every grid size in the campaign.
Control-plane cost differs: SDN Dijkstra recomputation remains on the order of
milliseconds, but docker-mediated route installation grows from roughly 3.5\,s to
14.2\,s at the first handover between 27 and 102 nodes, pushing 524 kernel routes
at $10\times 10$ versus 10\,504 routes at initialization.
OSPF route collection stays near 230\,ms but does not prevent the measured data-plane gaps.

Our contributions are:
(1)~a reproducible emulation methodology for graph-dynamic satellite networks with
symmetric control-plane and data-plane instrumentation;
(2)~evidence that proactive incremental SDN updates can eliminate handover black holes
up to 102-node topologies under the studied policy, at the expense of growing
install latency;
and (3)~a scaling dataset and analysis pipeline (batch experiments, aggregated CSVs,
and figure generation) released with the StarryNet artifact to support further
graph-centric routing research on dynamic infrastructure graphs.
\end{abstract}

\begin{IEEEkeywords}
Dynamic graphs,
software-defined networking,
satellite networks,
routing under topology change,
network emulation,
graph-centric systems,
LEO constellations,
data-plane availability,
incremental route installation,
scalability
\end{IEEEkeywords}
