% Experimental methodology section for OrbitGraph paper.

\section{Experimental Methodology}
\label{sec:experiments}

We evaluate OrbitGraph against a distributed OSPF baseline under identical
emulated constellations, traffic, and failure scripts.
All runs use StarryNet's full experiment profile on a single host; results are
aggregated across repeated, seeded trials and reported with mean and standard
deviation where applicable.

\subsection{Constellation scenarios and scaling grid}

We study six square grids from $5{\times}5$ through $10{\times}10$ orbital
planes and satellites per plane.
Each grid comprises $O \cdot S$ satellites plus two ground stations ($N = O
\cdot S + 2$ nodes total), ranging from 27 to 102 nodes.
Grids share the Starlink-inspired parameters in Section~\ref{sec:simulation}
(550\,km altitude, 53$^\circ$ inclination, plus-grid ISLs, LeastDelay policy,
instant GSL handover).
Ground-station endpoints for measurement are derived automatically as nodes
$O \cdot S + 1$ and $O \cdot S + 2$.

Emulation duration is sized per grid so that at least one GSL handover occurs
and a late steady-state sample is captured before teardown:
100\,s ($5{\times}5$), 120\,s ($6{\times}6$), 250\,s ($7{\times}7$),
400\,s ($8{\times}8$ and $10{\times}10$), and 350\,s ($9{\times}9$).
These values follow our paper-scale configuration and ensure handover-relative
probes fall inside the emulation window for every size.

\subsection{Compared routing modes}

We run two modes on each grid and seed.
In the OSPF baseline, the BIRD routing daemon runs OSPF in every container;
the emulation waits for OSPF convergence after initialization and reconverges
after each topology mutation.
In OrbitGraph (SDN), a centralized controller recomputes delay-weighted
shortest paths (Dijkstra) on each routing event, pushes only changed kernel
routes with make-before-break ordering (replace before delete), and performs
proactive handover: post-handover FIB entries are installed after new GSLs are
added but before old GSLs are removed.
All other emulation steps (delay updates every 10\,s, scripted link damage and
recovery, and ping scheduling) are identical between modes.
Only the control-plane path differs, yielding a paired comparison on the same
dynamic graph sequence.

\subsection{Workload and event script}

Each run follows the full profile.
At initialization, containers and links are created from second-1 delay matrices;
routing is installed (OSPF convergence or OrbitGraph initial FIB push), and a
ground-station-to-ground-station ping in the post-init phase verifies baseline
reachability.
At $t{=}5$, thirty percent of satellites are chosen pseudorandomly (seeded per
repetition); all interfaces on those nodes receive 100\% loss via traffic
control, modeling satellite failure.
At $t{=}10$, the same interfaces revert to the configured 1\% loss rate.
The first scheduled GSL handover adds a new ground--satellite link and removes
an old one at a geometry-dependent tick (e.g., $t{=}53$ on $5{\times}5$,
$t{=}23$ on $6{\times}6$).
Pings are scheduled at the handover instant, two seconds later
(post-handover), and near emulation end (steady), so phases align across grid
sizes despite different handover times.

\subsection{Metrics}

The primary data-plane outcome is outage duration around each routing event.
A continuous ICMP probe (10\,Hz) runs inside the source ground-station
container for the entire emulation; timestamps are aligned to control-plane
snapshot wall times.
Outage onset requires a sustained loss burst ($\geq$3 consecutive unreplied
ICMP sequences); recovery requires $\geq$3 consecutive replies.
This metric is symmetric for OSPF and OrbitGraph and captures multi-second OSPF
black holes that tick-scheduled pings would miss when route install blocks the
emulation loop.
We additionally report ping loss (\%) and mean RTT (ms) per phase (post-init,
handover, post-handover, steady) for the ground-station pair, and qualitative
path checks via traceroute where applicable.

For OrbitGraph we record graph computation time, kernel route install time,
nodes touched, and route counts (installed, deleted, failed) per snapshot
reason (init, damage recovery, topology change, proactive handover).
For OSPF we record BIRD route-dump collection time as an observation cost; it
is not equated with OrbitGraph install time.
Headline control-plane comparisons at handover contrast OSPF collection against
OrbitGraph compute plus incremental install at the proactive-handover snapshot.

\subsection{Repetitions and aggregation}

We execute ten independent repetitions per (grid, mode) pair (120 runs total
across the scaling study).
Repetition $i$ uses RNG seed $1000 + i$, fixing the damaged-satellite subset
while preserving natural container and kernel timing variation.
Runs are orchestrated by a batch driver with the full profile and merged into
summary tables for plotting.
Each batch checks for expected phases and routing-event reasons and flags
incomplete samples.
Figures in this paper aggregate the first-handover outage and handover-phase
loss/RTT unless noted otherwise; cross-size trends use $N = O \cdot S + 2$.

\subsection{Execution environment}

Experiments ran on a single Apple M2 Pro workstation (10 CPU cores, 32\,GiB
RAM) under macOS~14.6 with OrbStack providing the Linux container runtime
(Docker API 29.x, 19.6\,GiB allocated to the VM).
StarryNet local mode executes container and traffic-control commands on the
host without remote SSH.
Python~3.11 drives batch orchestration; each repetition tears down container
instances and associated bridge networks before the next seed starts.
This setup reflects a realistic researcher workstation deployment of
StarryNet~\cite{lai2023starrynet} and keeps OSPF and OrbitGraph runs colocated
on the same hardware, container image, and emulation timeline.
