% Conclusion section for OrbitGraph paper.

\section{Conclusion}
\label{sec:conclusion}

Low Earth orbit mega-constellations are, at their core, dynamic graphs:
vertices are satellites and ground assets, edges appear and vanish as orbital
geometry evolves, and edge weights track propagation delay and link capacity.
Routing in this environment is not a one-time shortest-path problem but a
continuous sequence of graph snapshots $G_t$ that must be reflected in the
forwarding plane faster than contacts change.
Distributed link-state protocols such as OSPF were designed for networks whose
topology changes sporadically; they discover mutations reactively and
reconverge after the fact, which our measurements show can black-hole traffic
for many seconds at each ground--satellite handover.

Software-defined networking offers a different contract.
By decoupling the control plane from the data plane, OrbitGraph
centralizes graph construction from ephemeris-driven delay
matrices and multi-source Dijkstra while leaving forwarding to the
existing Linux kernel on each node.
The data plane executes simple, fast operations (longest-prefix match, local
delivery); the control plane owns global visibility, policy, and when
updates land relative to link lifetime.
That separation is efficient in two senses: algorithmically, because
recomputation stays in the millisecond range even at 102 nodes; operationally,
because only changed next-hops are programmed (${\sim}10^2$–$10^3$ routes at
handover versus ${\sim}10^4$ at initialization), avoiding the cost of
re-flooding a full link-state database to every router on every orbital tick.

The decisive advantage, however, is not raw compute speed but control
over update ordering.
OrbitGraph's proactive handover installs the post-contact FIB while both old and
new ground--satellite links coexist, make-before-break at the topology level.
Distributed OSPF cannot match that behavior without protocol extensions,
because no single router observes the full mutation schedule in advance.
Our container-based evaluation on StarryNet~\cite{lai2023starrynet}, spanning
27 to 102 nodes with ten repetitions per configuration, shows 0\,ms mean
data-plane outage at the first GSL handover on every grid size versus
multi-second OSPF black holes (up to ${\sim}13.6$\,s), 0\% packet loss through
handover and post-handover phases under OrbitGraph while OSPF loss reaches 100\%
on several grids during reconvergence, comparable steady-state latency once both
modes settle, and incremental graph updates that keep programmatic cost
proportional to handover scope rather than constellation size.

OrbitGraph demonstrates that graph-centric SDN routing
can keep an emulated LEO network connected through predictable topology
changes, the regime where orbital mechanics already give operators a global
view and a schedule of contact events~\cite{kassing2020exploring,handley:ground-relays:2019}.
It is not a universal replacement for distributed routing: under unplanned bulk
failures with our container-based install path, centralized updates can lag OSPF
(Section~\ref{sec:analysis-damage}), and reaching thousands of nodes will require
hierarchical control and faster southbounds (Section~\ref{sec:future}).
Within its effective range (scheduled GSL handovers and orbit-predictable graph
snapshots), OrbitGraph shows that treating the constellation as an explicit,
time-indexed graph and programming forwarding from a centralized controller is
a practical path to the sub-second availability that latency-sensitive
space--terrestrial services require.

For the graph-centric computing perspective, when the
network is a graph that changes on a known timetable, the effective
computing model is to make that graph a first-class control object: recompute
on snapshots, push structured deltas, and align commits with graph edits,
rather than forcing each node to infer global state from local messages alone.
OrbitGraph is a concrete instantiation of that idea for LEO routing, validated
end-to-end in a real TCP/IP emulation environment with measurable gains in
data-plane availability where it matters most.
